---
title: "Wooldridge Computer Exercise Ch16.C1"
output: 
  pdf_document:
    latex_engine: xelatex
fontsize: 12pt
header-includes:
 - \setmainfont{Adobe Garamond Pro}
 - \setmonofont[Scale=MatchLowercase]{Source Code Pro}
 - \setsansfont[Scale=MatchLowercase]{Myriad Pro}
---


```{r}
library(wooldridge)
data("smoke")
head(smoke)
```
## i)
$\beta_1$ is the change in $\log(\texttt{income})$ for an increase of one cigarette a day.

## ii) 
We would expect a decrease in consumption for an increase in cigarette prices, therefore $\gamma_5<0$, and we would also expect cigarette consumption to decrease if there is a ban on smoking in restaurants, $\gamma_6<0$. 

## iii) 
If $\gamma_5\neq 0 and \gamma_6\neq 0$.

## iv) 
```{r}
olsinc <- lm(log(income)~cigs+educ+age+I(age^2), data=smoke)
summary(olsinc)
```
$\beta_1$ is slightly positive, which means that income slightly increases with an increase in smoking, but it is also not statistically significant.

## v)
```{r}
reducedform <- lm(cigs ~ educ+age+I(age^2)+log(cigpric)+restaurn, data=smoke)
summary(reducedform)
```
restaurn is significant, but $\log(cigpric)$ is not. We can still estimate the regression, as we have one IV left. 

## vi)
```{r}
smoke$fitcigs <- fitted(reducedform)
secondstage <- lm(log(income)~fitcigs+educ+age+I(age^2), data=smoke)
summary(secondstage)
```
We now get the expected sign for $\beta_1$.

# Using R functions
```{r}
library(AER)
iv <- ivreg(log(income)~cigs+educ+age+I(age^2)|educ+age+I(age^2)+log(cigpric)+restaurn, data=smoke)

summary(iv,vcov = sandwich, diagnostics = TRUE)
#se slightly off, but acceptable.
#--------
#not really working
library(systemfit)
system <- systemfit(formula=list(log(income)~cigs+educ+age+I(age^2),cigs~log(income)+educ+age+I(age^2)+log(cigpric)+restaurn), 
          inst=~log(cigpric)+restaurn, 
          method="2SLS",
          data=smoke)
summary(system)
```